$\pi$ with leftovers: \\ a mechanisation in Agda

### 2020-05-06 paper/paper.tex
No need for parentheses around tuples

  Henceforth, we use $\lz$ to denote the pair $(0 \comma 0)$ for a channel which cannot be used any further, $\li$ for the pair $(1 \comma 0)$ for a channel to be used in input exactly once, $\lo$ for $(0 \comma 1)$ for a channel to be used in output exactly once, and $\lio$ for $(1 \comma 1)$, for a channel to be used exactly once in input and exactly once in output.
  This notation was originally used in the linear \picalc{} \cite{KPT96,Sangio01}.
\end{note}

Linear, graded and shared types are all defined as an instance of our usage algebra.
Their use in typing derivations is illustrated in \autoref{example-derivations}.
\begin{description}
  \item [Linear]
    The carrier is a type with two trivial constructors $\constr{zero}$ and $\constr{one}$.
    The monoidal operation has the element $\constr{zero}$ as neutral on both sides, and the element $\constr{one}$ splitting into $\constr{one}$ and $\constr{zero}$ (or $\constr{zero}$ and $\constr{one}$), and is uninhabited otherwise.
    All the properties of the algebra follow trivially.
    As an example, the multiplicity pair $\ell_i$ is represented as $(\constr{one} \comma \constr{zero})$.

    \item [Graded]
    The carrier is the type of natural numbers.
    The element $\zero$ corresponds to 0, and $\one$ to 1.
    The partial monoid $\op{x}{y}{z}$ is defined exactly when $x = y + z$.
    All other properties of the algebra follow from the algebraic rules for the addition of natural numbers.
    As an example, we use $(5 \comma 3)$ to represent the multiplicities of a channel that is able to input 5 times and output 3 times.

    \item [Shared]
    The carrier is a type with a single trivial constructor $\constr{\omega}$.
    Both $\zero$ and $\one$ are instantiated to $\constr{\omega}$.
    The relation $\op{\constr{\omega}}{\constr{\omega}}{\constr{\omega}}$ is inhabited.
    All the properties of the algebra follow trivially.
    Shared channels have multiplicities $(\constr{\omega} \comma \constr{\omega})$, which they will forever preserve.
\end{description}

@@

  The inaction process in rule $\constr{end}$ does not change usage annotations.
  The scope restriction in rule $\constr{chan}$ expects three arguments: the type $t$ of data being transmitted; the usage annotation $x$ of what is being transmitted; and the multiplicity $y$ given to the channel itself.
  This multiplicity $y$ is used for both input and output, so that they are balanced.
  The continuation process $P$ is provided with the new channel with usage annotation $(y \comma y)$, which it must completely exhaust.

@@

  Let $\func{\omega}$ be a shortcut for $(\one \comma \one)$ in an algebra of shared types.
  Let $\func{\epsilon}$ be the empty usage context --- where all usage annotations are $(\zero \comma \zero)$.
  To type $\func{p}$, we need a channel over which we can transmit another channel.
  We therefore create a new channel with usage annotations $(\one \comma \one)$ capable of transmitting data of type $\channel{\unit}{\func{\omega}}$, and declare (with $\li$) that we only transmit input capabilities over it (line 7).
  The left-hand process is typed following its syntax (line 8).
  On the right-hand we need the channel that is going to be sent (line 9).
  The type $\unit$ of this channel is entirely determined by the type $\channel{\unit}{\func{\omega}}$ of the channel over which it is going to be transmitted.
  Its usage annotation is $(\one \comma \one)$: we will transmit $(\one \comma \zero)$ and save $(\zero \comma \one)$ for the continuation process.

>>>>

  Henceforth, we use $\lz$ to denote the pair $0 \comma 0$ for a channel which cannot be used any further, $\li$ for the pair $1 \comma 0$ for a channel to be used in input exactly once, $\lo$ for $0 \comma 1$ for a channel to be used in output exactly once, and $\lio$ for $1 \comma 1$, for a channel to be used exactly once in input and exactly once in output.
  This notation was originally used in the linear \picalc{} \cite{KPT96,Sangio01}.
\end{note}

Linear, graded and shared types are all defined as an instance of our usage algebra.
Their use in typing derivations is illustrated in \autoref{example-derivations}.
\begin{description}
  \item [Linear]
    The carrier is a type with two trivial constructors $\constr{zero}$ and $\constr{one}$.
    The monoidal operation has the element $\constr{zero}$ as neutral on both sides, and the element $\constr{one}$ splitting into $\constr{one}$ and $\constr{zero}$ (or $\constr{zero}$ and $\constr{one}$), and is uninhabited otherwise.
    All the properties of the algebra follow trivially.
    As an example, the multiplicity pair $\ell_i$ is represented as $\constr{one} \comma \constr{zero}$.

    \item [Graded]
    The carrier is the type of natural numbers.
    The element $\zero$ corresponds to 0, and $\one$ to 1.
    The partial monoid $\op{x}{y}{z}$ is defined exactly when $x = y + z$.
    All other properties of the algebra follow from the algebraic rules for the addition of natural numbers.
    As an example, we use $5 \comma 3$ to represent the multiplicities of a channel that is able to input 5 times and output 3 times.

    \item [Shared]
    The carrier is a type with a single trivial constructor $\constr{\omega}$.
    Both $\zero$ and $\one$ are instantiated to $\constr{\omega}$.
    The relation $\op{\constr{\omega}}{\constr{\omega}}{\constr{\omega}}$ is inhabited.
    All the properties of the algebra follow trivially.
    Shared channels have multiplicities $\constr{\omega} \comma \constr{\omega}$, which they will forever preserve.
\end{description}

@@

  The inaction process in rule $\constr{end}$ does not change usage annotations.
  The scope restriction in rule $\constr{chan}$ expects three arguments: the type $t$ of data being transmitted; the usage annotation $x$ of what is being transmitted; and the multiplicity $y$ given to the channel itself.
  This multiplicity $y$ is used for both input and output, so that they are balanced.
  The continuation process $P$ is provided with the new channel with usage annotation $y \comma y$, which it must completely exhaust.

@@

  Let $\func{\omega}$ be a shortcut for $\one \comma \one$ in an algebra of shared types.
  Let $\func{\epsilon}$ be the empty usage context --- where all usage annotations are $\zero \comma \zero$.
  To type $\func{p}$, we need a channel over which we can transmit another channel.
  We therefore create a new channel with usage annotations $\one \comma \one$ capable of transmitting data of type $\channel{\unit}{\func{\omega}}$, and declare (with $\li$) that we only transmit input capabilities over it (line 7).
  The left-hand process is typed following its syntax (line 8).
  On the right-hand we need the channel that is going to be sent (line 9).
  The type $\unit$ of this channel is entirely determined by the type $\channel{\unit}{\func{\omega}}$ of the channel over which it is going to be transmitted.
  Its usage annotation is $\one \comma \one$: we will transmit $\one \comma \zero$ and save $\zero \comma \one$ for the continuation process.
